---
title: "Assisgnment 2"
author: "Suyashkumar Chavan(228001537)"
date: "10/5/2018"
output:
  word_document: default
  pdf_document: default
---
```{r}

```

#Problem 1
Assignment #2: Linear Regression
Use the Auto data set to answer the following questions:
##(a) Perform a simple linear regression with mpg as the response and horsepower as the predictor.
```{r}
library(ISLR)
data(Auto)
Relation <- lm(mpg~horsepower,data=Auto)
summary(Relation)
```
##i. Is there a relationship between the predictor and the response?
Ans:The p-value corresponding to F-statistic is very low(<0.001).So it indicates that there is a relationship between response amd predictor.

##ii. How strong is the relationship between the predictor and the response?
Ans:The R-squared value is 0.6059, which means 60.59% of the variation in mpg can be explained using horsepower data.

##iii. Is the relationship between the predictor and the response positive or negative?
Ans: The relationship between mpg and horsepower is negative, as co-efficient of horsepower is negative i.e. -0.157845. It shows that with increase in horsepower, there is reduction in mpg.

##iv. How to interpret the estimate of the slope?
Ans: Coefficient of Horsepower is -0.157845, so with increase of 100 in horsepower, mpg will decrease by 15.7845.

##v. What is the predicted mpg associated with a horsepower of 98? What are the associated 95% confidence and prediction intervals?

```{r}
predict(Relation, data.frame(horsepower=98),interval="confidence")

```

```{r}
predict(Relation, data.frame(horsepower=98), interval="prediction")
```

##(b) Plot the response and the predictor. Display the least squares regression line in the plot.
```{r}
plot(Auto$horsepower,Auto$mpg,main="plot of mpg vs horsepower",xlab="horsepower",ylab="mpg",col="red")
abline(Relation,col="Blue")
```


##(c) Produce the diagnostic plots of the least squares regression fit. Comment on each plot.

```{r}
par(mfrow=c(2,2))
plot(Relation)
```

Comment 1: From the plot of residuals vs Fitted values, The non-linearity relationship is present in the data.
Comment2: From the Residuals vs Leverage plot, there are many high leverage points present in the data and also the the outliers.
Comment3:Normal Q-Q Plot: From the Normal Q-Q plot, it is a fairly straight line, so both set of quantiles is come from normal distribution.
comment4:Scale-location plot:From scale-location plot, residuals are spread randomly.



##(d) Try a few different transformations of the predictor, such as log(ð‘‹) , âˆšð‘‹, ð‘‹2, and repeat (a)-(c). Comment on your findings.


Ans:
1.log(x)
##(a) Perform a simple linear regression with mpg as the response and log(horsepower) as the predictor.
```{r}
Relation2 <- lm(mpg~log(horsepower),data=Auto)
summary(Relation2)
```

##i. Is there a relationship between the predictor and the response?
Ans:The p-value corresponding to F-statistic is very low(<0.001).So it indicates that there is a relationship between response amd predictor.

##ii. How strong is the relationship between the predictor and the response?
Ans:The R-squared value is 0.6683, which means 66.83% of the variation in mpg can be explained using horsepower data.

##iii. Is the relationship between the predictor and the response positive or negative?
Ans: The relationship between mpg and horsepower is negative, as co-efficient of log(horsepower )is negative i.e. -18.5822. It shows that with increase in horsepower, there is reduction in mpg.

##iv. How to interpret the estimate of the slope?
Ans: Coefficient of log(Horsepower) is -18.5822, so with increase of 1 in log(horsepower), mpg will decrease by 18.5822.

##v. What is the predicted mpg associated with a horsepower of 98? What are the associated 95% confidence and prediction intervals?
```{r}
predict(Relation2, data.frame(horsepower=98),interval="confidence")
```

```{r}
predict(Relation2, data.frame(horsepower=98), interval="prediction")
```

##(b) Plot the response and the predictor. Display the least squares regression line in the plot.


```{r}
plot(log(Auto$horsepower),Auto$mpg,main="plot of mpg vs horsepower",xlab="log(horsepower)",ylab="mpg",col="red")
abline(Relation2,col="Blue")
```


##(c) Produce the diagnostic plots of the least squares regression fit. Comment on each plot.
Ans:
```{r}
par(mfrow=c(2,2))
plot(Relation2)
```

Comment1:in residual vs fitted plot, wih log of horsepower, line is fairly straight, so it justifies non-linearity.
Comment2:in residual vs leverage plot,From the Residuals vs Leverage plot, there are many high leverage points present in the data and also the the outliers.
Comment3:Normal Q-Q Plot: From the Normal Q-Q plot, it is a fairly straight line, so both set of quantiles is come from normal distribution.
comment4:Scale-location plot:From scale-location plot, residuals are spread randomly.

2.sqrt(x)
##(a) Perform a simple linear regression with mpg as the response and sqrt(horsepower) as the predictor.
```{r}
Relation3 <- lm(mpg~sqrt(horsepower),data=Auto)
summary(Relation3)
```
##i. Is there a relationship between the predictor and the response?
Ans:The p-value corresponding to F-statistic is very low(<0.001).So it indicates that there is a relationship between response amd predictor.

##ii. How strong is the relationship between the predictor and the response?
Ans:The R-squared value is 0.6437, which means 64.37% of the variation in mpg can be explained using horsepower data.

##iii. Is the relationship between the predictor and the response positive or negative?
Ans: The relationship between mpg and horsepower is negative, as co-efficient of log(horsepower )is negative i.e. -3.503. It shows that with increase in squareroot of horsepower, there is reduction in mpg.

##iv. How to interpret the estimate of the slope?
Ans: Coefficient of sqrt(Horsepower) is -3.503, so with increase of 1 in sqrt(horsepower), mpg will decrease by 3.503.

##v. What is the predicted mpg associated with a horsepower of 98? What are the associated 95% confidence and prediction intervals?
```{r}
predict(Relation3, data.frame(horsepower=98),interval="confidence")
```

```{r}
predict(Relation3, data.frame(horsepower=98),interval="prediction")
```

##(b) Plot the response and the predictor. Display the least squares regression line in the plot.
```{r}

plot(sqrt(Auto$horsepower),Auto$mpg,main="plot of mpg vs horsepower",xlab="sqrt(horsepower)",ylab="mpg",col="red")
abline(Relation3,col="Blue")

```
##(c) Produce the diagnostic plots of the least squares regression fit. Comment on each plot.

```{r}
par(mfrow=c(2,2))
plot(Relation3)
```


Comment1:in residual vs fitted plot, wih log of horsepower, line is fairly straight, so it justifies non-linearity.
Comment2:in residual vs leverage plot,From the Residuals vs Leverage plot, there are many high leverage points present in the data and also the the outliers.
Comment3:Normal Q-Q Plot: From the Normal Q-Q plot, it is a fairly straight line, so both set of quantiles is come from normal distribution.
comment4:Scale-location plot:From scale-location plot, residuals are spread randomly.

3.square(x)
##(a) Perform a simple linear regression with mpg as the response and square(horsepower) as the predictor.
Ans:
```{r}
Relation4 <- lm(mpg~(horsepower)^2,data=Auto)
summary(Relation4)
```
##i. Is there a relationship between the predictor and the response?
Ans:The p-value corresponding to F-statistic is very low(<0.001).So it indicates that there is a relationship between response amd predictor.

##ii. How strong is the relationship between the predictor and the response?
Ans:The R-squared value is 0.6059, which means 60.59% of the variation in mpg can be explained using horsepower data.

##iii. Is the relationship between the predictor and the response positive or negative?
Ans: The relationship between mpg and horsepower is negative, as co-efficient of horsepower is negative i.e. -0.157845. It shows that with increase in horsepower, there is reduction in mpg.

##iv. How to interpret the estimate of the slope?
Ans: Coefficient of Horsepower is -0.157845, so with increase of 100 in horsepower, mpg will decrease by 15.7845.

##v. What is the predicted mpg associated with a horsepower of 98? What are the associated 95% confidence and prediction intervals?

```{r}
predict(Relation4, data.frame(horsepower=98),interval="confidence")
```


```{r}
predict(Relation4, data.frame(horsepower=98),interval="prediction")
```

##(b) Plot the response and the predictor. Display the least squares regression line in the plot.
```{r}
plot((Auto$horsepower)^2,Auto$mpg,main="plot of mpg vs horsepower",xlab="square(horsepower)",ylab="mpg",col="red")
abline(Relation4,col="Blue")
```

##(c) Produce the diagnostic plots of the least squares regression fit. Comment on each plot.
```{r}
par(mfrow=c(2,2))
plot(Relation4)
```
Ans:
Comment 1: From the plot of residuals vs Fitted values, The non-linearity relationship is present in the data.
Comment2: From the Residuals vs Leverage plot, there are many high leverage points present in the data and also the the outliers.
Comment3:Normal Q-Q Plot: From the Normal Q-Q plot, it is a fairly straight line, so both set of quantiles is come from normal distribution.
comment4:Scale-location plot:From scale-location plot, residuals are spread randomly.




#Problem2
Use the Auto data set to answer the following questions:
(a) Produce a scatterplot matrix which includes all of the variables in the data set. Which predictors
appear to have an association with the response?
```{r}
pairs(Auto)
```


Association between response and predictor:
All the predictors are associated with reponse except horsepower and acceleration.

##(b) Compute the matrix of correlations between the variables (using the function cor()). You will
need to exclude the name variable, which is qualitative.
Ans:
```{r}
names(Auto)
```

```{r}
cor(Auto[1:8])
```


##(c) Perform a multiple linear regression with mpg as the response and all other variables except name as the predictors.
```{r}
relation5<- lm(mpg~.-name,data=Auto)
summary(relation5)
```

##i. Is there a relationship between the predictors and the response?
Ans: The p-value related to the given F-statitstic is very low(<0.01), So there is relationship between mpg and other predictors.

##ii. Which predictors have a statistically significant relationship to the response?
Ans:All the predictors have significant relationship with the response except the cylinders, horsepower and acceleration, as their p-value related to t-statistics are fairly high.

##iii. What does the coefficient for the year variable suggest?
Ans: When all other predictors are kept constant, then with increase in 1 year of car, its mpg will increase ny 0.750773. Hence there is a positive relationship.

##(d) Produce diagnostic plots of the linear regression fit. Comment on each plot.
Ans:
```{r}
par(mfrow=c(2,2))
plot(relation5)
```
Comment 1: From the plot of residuals vs Fitted values,as red line is fairly straight, linearity relationship is present in the data.
Comment2: From the Residuals vs Leverage plot, there are many high leverage points present in the data and also the the outliers.
Comment3:Normal Q-Q Plot: From the Normal Q-Q plot, it is a fairly straight line, so both set of quantiles is come from normal distribution.
comment4:Scale-location plot:From scale-location plot, residuals are spread randomly.

##(e) Is there serious collinearity problem in the model? Which predictors are collinear?

```{r}
car::vif(relation5)
```
vif value of cylinders, displacement,horsepower,weight,acceleration is high, hece these predictors are collinear.




#Problem3
Use the Carseats data set to answer the following questions:

#(a) Fit a multiple regression model to predict Sales using Price, Urban, and US.
```{r}
data(Carseats)
relation6 <- lm(Sales~Price+Urban+US,data=Carseats)
summary(relation6)
```

##(b) Provide an interpretation of each coefficient in the model (note: some of the variables are qualitative).
Ans:
1.Price:With average increase in price by $1, there is decrease of 54.459(average)units of sales.
2.Urban:Keeping the other predictors constant, sales in urban area are less than rural area by 21.916 units.
3.US:keeping the other predictors constant , sales in US are more than Non-USstores by 1200.573(average)units.

##(c) Write out the model in equation form.
Ans.
    Sales= 13.043469+ (-0.054459) X Price + (-0.021916) X Urban + (1.200573) X US + e
    where, e=error.
    
##(d) For which of the predictors can you reject the null hypothesis ð» : ð›½ = 0 ?
Ans: Null Hypothesis can be rejected for "Price" and "US" variables, as p-value for these predictors is fairly low.

##(e) On the basis of your answer to the previous question, fit a smaller model that only uses the predictors for which there is evidence of association with the response.
Ans:
```{r}
data(Carseats)
relation7 <- lm(Sales~Price+US,data=Carseats)
summary(relation7)
```

##(f) How well do the models in (a) and (e) fit the data?
Ans:
The R^2 value for new model (e) is better than older model(a). more variation is explained in the newer model.

##(g) Is there evidence of outliers or high leverage observations in the model from (e)?
Ans:
```{r}
par(mfrow=c(2,2))
plot(relation7)
```

there are some outliers are present in the model, because in residual vs leverage plot, many points are outside of -2 to 2.
There are leverage points are also present.




